---
title: "Comorbidity package"
author: "David Henderson"
date: "24 August 2018"
output: 
  html_document:
    theme: spacelab
    code_folding: hide
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


#Introduction

A quick play with the `comorbidity` package to see the difference in Charlson and Elixhauser scores when using full and truncated ICD10 codes. 

##Packages

```{r lib}
library(comorbidity)
library(tidyverse)
library(magrittr)
```

##Initialise dataframes

The package lets you make dummy data really easily. I'm going to make 30,000 observations of 5000 indiviudals (differing numbers of disease) then create a copy dataset with ICD10 codes truncated to 3 characters.

```{r df}
#Full data
data_full_code <- data.frame(
  id = sample(1:5000, size = 30000, replace = TRUE),
  code = sample_diag(n = 30000, version = "ICD10_2011"),
  stringsAsFactors = FALSE
)

data_full_code %<>%
  arrange(id)
#Truncated data
data_truncated <-
  data_full_code %>% 
  mutate(code = strtrim(code, 3))

data_full_code
data_truncated
```

#Calculate Charlson Index

Now run the `comorbidity()` function to create Charlson scores for both datasets.

```{r ch}
charlson_full <- comorbidity(data_full_code, id = "id", code = "code", score = "charlson_icd10")
charlson_trunc <- comorbidity(data_truncated, id = "id", code = "code", score = "charlson_icd10")
charlson_full
charlson_trunc
```

Are they the same

```{r check}
identical(charlson_full, charlson_trunc)
```


Of course not, how different?

##Visualise

Scores for full data

```{r viz}
charlson_full %>% 
  ggplot(aes(score)) +
  geom_bar()
```

...and truncated data. Slight drops in higher scores and increases in lower/zero scores as expected but not a huge difference

```{r viz_2}
charlson_trunc %>% 
  ggplot(aes(score)) +
  geom_bar()
```

The `comorbidity()` function also calculates a weighted score which increases scores for certain diseases (e.g. diabetes with complications). I'd expect to see a bigger difference here...

```{r we_ch}
charlson_full %>% 
  ggplot(aes(wscore)) +
  geom_bar()
```

truncated data...

```{r trunc_weight}
charlson_trunc %>% 
  ggplot(aes(wscore)) +
  geom_bar()
```

This holds up surprisingly well...

#Elixhauser

Repeat for elixhauser

```{r elix}
elix_full <- comorbidity(data_full_code, id = "id", code = "code", score = "elixhauser_icd10")
elix_trunc <- comorbidity(data_truncated, id = "id", code = "code", score = "elixhauser_icd10")
elix_full
elix_trunc
```

##Visualise

Full codes

```{r elix_1}
elix_full %>% 
  ggplot(aes(score)) +
  geom_bar()
```

Truncated

```{r trunc}
elix_trunc %>% 
  ggplot(aes(score)) +
  geom_bar()
```

Again, the same pattern where there are more zero and 1 scores and fewer higher scores. Again this is a very mild change though. 

Weighted score full data

```{r elix_wh}
elix_full %>% 
  ggplot(aes(wscore)) +
  geom_bar()
```

truncated

```{r trunc_elix_we}
elix_trunc %>% 
  ggplot(aes(wscore)) +
  geom_bar()
```

Again holds up better than expected.

Of course - this is dummy data and difficult to know how well this would work in the real world!

#session_info

```{r session}
devtools::session_info()
```